Have you forgotten?' Conversion experiences in shopping lists

### 2017-02-17 lists_conversion_experience.rmd
tidying up; adding bit of Miller

---
title: "'Have you forgotten?' Conversion experiences in shopping lists"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: /home/mackenza/ref_bibs/uni.bib
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

@@

```{r setup, cache=FALSE, include=FALSE}
library(knitr)
library(tint)
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
output <- opts_knit$get("rmarkdown.pandoc.to")
opts_chunk$set(warning=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(echo=FALSE)
opts_chunk$set(error=FALSE)
opts_chunk$set(cache=TRUE)
```

@@

This paper investigates practices of list-making associated with online grocery shopping at a UK-global supermarket chain, Tesco. 
It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
The recommender system  aim to up-sell and cross-sell some of the 250,000 items in the supermarket's inventory. 
I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
While all written lists mix past and future actions, actual and potential events, their re-writing by recommender systems adds a new operational dynamic.
The incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.

@@

Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'. When customers shop for groceries online, a list of five recommendations appear at the checkout stage. These recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).  The question 'have you forgotten?' is followed by some  grocery items that could have been or are usually on a shopping list. The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know what should be on your list, perhaps because you have never thought of putting them there.

The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).  Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital. Encounters between this hyper-complex geography and people, even in the confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco. While there are many facets to shopping [@Miller_2013], the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative). As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.

While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists take shape at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems. Just as the aisles and shelves of a supermarket present a commercially-ordered geography of the world,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.

@@

Despite being titled 'Have You Forgotten?' Tesco's recommender system does not suggest forgotten items.  It focuses on 'conversion events.' If a shopper does buy  recommended items, 'conversion' has occurred.  From a supermarket perspective, high-value conversion are desirable. There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.  From a data science perspective, conversion events, when they occur, confirm the power of predictive analytics to change the world in some way. From a sociological perspective, matters are less obvious. Constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data. We might regard the recommender system as an operational formation that responds to an ongoing control crisis associated with information [@Beniger_1986]. We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what the philosopher science Karl Popper terms the propensities of things [@Popper_1990, 11] -- happens in and through data practices.   Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.  Recommender systems, then, are vitally concerned not just with conversion events, but also with the what counts as order in the world.

## Lists and order

Lists have long interested historians, anthropologists, sociologists, literary and media theorists. If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g. tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices. As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance. More recent sociological work on lists (see [@deGoede_2016] for an oveview) confirms the potency of lists as ordering devices. From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally. Lists such as online grocery recommendations weave together people, infrastructures, things, places and predictions.[^1.456]

@@

Hansen predicates prediction as an 'ontological transformation' thata ffects probabilities. It links probabilities as calculations to propensities or the mutable agencies of things. The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  recommendations. I will argue that 'conversion' is a preferable term for these changes.  It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's subjective 'understood-to-be-expressed' formulation.

@@

In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive modles.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/). Her work focuses on online grocery shopping at the supermarket chain Tesco.  Speaking to an audience of statisticians and operations researchers, Patel   focused on the models underlying  shopping list recommendations. Her presentation was filled with graphs, numbers, and tables concerning different aspects of the 'Have you forgotten?' recommender system. Patel's presentation was not sociological. In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])  A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries) and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].[^11.2]

@@

[^11.2]:

@@

Figure \@ref(fig:simulate-inventory) plots something of the density of commodities in the Tesco grocery database. By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10.

@@

I respond to Patel's presentation in two ways. First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in many commercial settings. Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice.  Second, by trying to step into the list-making practices  as an occasion to apprehend in the wider conversions they signal. As a sociological response to a data scientist presenting a model to an audience of statisticians and operations researchers, my interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.  The method of reconstruction borrows from the philosopher John Dewey a broad interest in how practical and abstract elements of experience are interwoven [@Dewey_1957]. More concretely, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.

@@

The main narrative of Patel's presentation concerned a shift from a well-established demographically-targeted data mining model for making recommendations to a predictive, probabilistic, personally relevant  re-writing of the shopping list in almost-realtime. The changes Patel described are increasingly widespread. While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly symptomatic. Tesco is the largest supermarket chain in UK, famous for its customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991. DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is famous for having convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.  Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984]. The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer preferences.

@@

```{r arules, cache=TRUE, messages=FALSE, echo=FALSE, fig.margin=TRUE}
library(knitr)
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
options(digits=2)
rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=F))
rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
kable(inspect(rules_sorted[1:5]))
```

The `apriori` algorithm finds sets of items that commonly occur together in transactions.  These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules) shows how often common items appear. `Whole milk` appears most frequently.  The `apriori` algorithm attempts to re-structure shopping lists using associations between items frequently bought together.  The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]). The association between items is expressed in the form of _rules_ not a model. These rules, a term derived from older decision support literature, define sets of associations between items. As the product of a  counting practice, the association rules define sets of items or shopping baskets in terms of decreasing strength. Table \@ref(tab:arules) shows that the first rule has a stronger association than the second. Milk and vegetables is more likely an association than milk and buns.  A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.

@@

```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
    library(ggplot2)
    tesco = read.csv('data/reference_data/tesco_groceries.csv')
    total = sum(tesco$actual, na.rm=TRUE) 
    tesco = na.omit(tesco)
    tesco$actual = as.integer(tesco$actual)

    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
```

Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations. The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items.  This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in, or very close to, the recommender system rather than  'in the world.' The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered. The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial. `r total` products (actually Patel mentioned 200,000 products) can be combined ways. If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists. Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially highly interesting and value as recommendation.

@@

```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
library(ggplot2)
library(reshape2)
p = as.logical(rnorm(n=2e5, size=1, prob=0.01))
m = (matrix(p, nrow=1000, byrow=TRUE))
melted <- melt(m)
melted$value <- as.logical(melted$value)
ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
```

@@

## Infrastructure: realtime engine serves up top 5

@@

In their intersection with mundane practices of ordering, recommender systems offer some guidance to the ways that predictive practices affect existing  forms of social order in the world. Rather than directly positing an ontological transformation, I have explored some of the practices of conversion associated with a shift from a data-mining rule-based system to a predictive recommender system. The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data. We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time. The predictive assemblages respond to problems of logistical ensembles more generally.

Following Hansen, I have been suggesting that the predictive listing in the form of open-ended recommendation changes social order. It alters the way in which repetition or reiteration, exemplified in the list, participating in real situations, and their possibilities. By following Popper in understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.

I diverge from Hansen in making sense of predictive practice. Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggest that predictive conversion remains constitutively incomplete. Predictive practice abounds in propensities. Ironically, shopping lists  exemplify this incompleteness and ineradicable openness of prediction practice.  The constitutive incompleteness and openness of lists are very closely entangled with problems of social order amidst predictive operations. Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset. The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiment with structures that lack closure. A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion event.

@@

The 'universe of reference' includes objects, ideas and practices mingle. Something similar happens in the Dunnhumby relevance model. In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape. The shopping list, as a result, take on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities. Shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.

>>>>

---
title: "'Have you forgotten?' Conversion experiences in shopping lists"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: /home/mackenza/ref_bibs/uni.bib
bibliostyle: 'harvard'
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

@@

```{r setup2, cache=FALSE, include=FALSE}
library(knitr)
library(tint)
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
output <- opts_knit$get("rmarkdown.pandoc.to")
opts_chunk$set(warning=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(echo=FALSE)
opts_chunk$set(error=FALSE)
opts_chunk$set(cache=TRUE)
```

@@

This paper investigates practices of list-making associated with online grocery shopping at a UK-global supermarket chain, Tesco. 
It reconstructs an account of how a recommender system suggests a small number of grocery items of personal relevance to each of the millions of online grocery shoppers. 
I suggest that we can understand devices such as recommender systems as transformations or conversions of the temporal and agential fabric of the important social ordering device, the list.
While all written lists mix past and future actions, actual and potential events, their re-writing by recommender systems adds a new operational dynamic.
The incompleteness of lists, their propensity to expand or change, becomes more important than their capacity to  enumerate or rank things. 
The conversion of lists from finite to in-definite in recommender systems suggests we might need to re-conceptualise how we think about social order and structure in media cultures.

@@

Online grocery shopping at Tesco, a large supermarket chain based in the UK, includes recommendations for further grocery purchases under the title of 'Have you forgotten?'. When customers shop for groceries online, a list of five recommendations appear at the checkout stage. These recommendations are the product of a recommender system, an important category of operational device in big data (see for instance [@Striphas_2015] for analysis of Netflix recommendation; [@Morris_2015]  or [@Seaver_2015] for an account of music recommendations]).  The question 'have you forgotten?' is followed by a list of some  grocery items that could have been or are usually on a shopping list. The title of the suggestions is a bit misleading: the recommender system, as we will see, is not concerned with forgetting, with the many slips and oversights associated with shopping lists,  but rather with items that you didn't know should be on your list, perhaps because you have never thought of putting them there.

The shelves of large contemporary supermarket are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity (see [@Neilson_2012; @Tsing_2009] on logistics and power are the endpoint of global logistic supply chains, in all their logistical, value-transforming and brand-mediated hypercomplexity).  Groceries imply a geography of agriculture, industry, transport, communication and financialisation driven by flows of labour and capital. Encounters between this hyper-complex geography and people, even in the familiar confines of a supermarket, are no simple matter, either for shoppers or for supermarket operators such as Tesco. While there are many facets to shopping (see [@Miller_2012] for an account of shopping as a practice that addresses discrepancies between normative and actual social order, and as a sacrificial ritual that instates social order), the shopping list, whether written on the back of an envelope, or saved as a list in an online grocery shopping system, is one intersectional ordering device that configures a universe of possible references, and a teeming multitude of propensities, into a possible order (in both senses of that term: an ordering and an imperative). As people shop, either by  trawling along aisles packed with thousands of products, or scrolling down screens or searching for particular brands amidst search results, lists inscribe some kind of order that filters or reduces the excessive propensities, claims, dazzle and distraction of commodities.

While hand-written shopping lists undoubtedly have ongoing importance and display interestingly mixed ordering practices (see the list of handwritten shopping lists at [Grocery List](http://www.grocerylist.org)), online shopping lists are written and re-written at the intersection of web and internet infrastructures, supply chain logistics, individualised practices and *habitus,* and increasingly, the predictive operations of  recommender systems. Just as the aisles and shelves of a supermarket present a commercially-ordered geography of needs and desires,  online shopping recommender systems generate lists that seek to align people to commodities they otherwise might have little relation to.

@@

Despite being titled 'Have You Forgotten?' Tesco's recommender system does not suggest forgotten items.  It focuses on 'conversion events.' If a shopper does buy  recommended items, 'conversion' has occurred.  From a supermarket perspective, high-value conversion are desirable. There is little value in successfully recommending small or cheap items such as rubber bands or toothpicks.  From a data science perspective, conversion events, when they occur, confirm the power of predictive analytics to change the world in some way. From a sociological perspective, matters are less obvious. Constructing and configuring recommender system in the name of conversion events, the recommender system effects an epistemic-operational transformation central to big data. We might regard the recommender system as an operational formation that responds to an ongoing control crisis associated with information [@Beniger_1986]. We might also explore these conversion events in terms of how predictive analytics happens in the world, how the tendency of things to realize certain events -- what Popper terms the 'propensities of things' [@Popper_1990, 11] -- happens in and through data practices.   Data in concert with calculation, models, infrastructures, and experiments, affords apprehension, visibility and management of associative propensities in the world.  Recommender systems, then, are vitally concerned not just with conversion events, but also with the what counts as order in the world.

## Lists and orders

Lists have long interested historians, anthropologists, sociologists, literary and media theorists. If, as the social anthropologist Jack Goody argued, lists are historically primary as forms of writing [@Goody_1986], and if list-making and its later variants (e.g. tables) precede discursive and narrative writing practices, then we might expect lists to function as powerful social ordering devices. As literary scholars suggest (see [@Mainberger_2003]), although lists have often been de-valued as literary forms, list-making often appears, even in literary form, whenever writing seeks to address, name, group or evoke totality, profusion, excess or abundance. More recent sociological work on lists (see [@deGoede_2016] for an overview) confirms the potency of lists as ordering devices. From both  anthropological and literary perspectives, shopping lists have excellent reasons to exist: they are deeply rooted in urban social formations historically and culturally. Lists such as online grocery recommendations weave together people, infrastructures, things, places and predictions.[^1.456]

@@

Hansen predicates prediction as an 'ontological transformation' that affects probabilities. It links probabilities as calculations to propensities or the mutable agencies of things. The practical and indeed empirical  question is whether such transformations in probability that bring 'the propensities of things'  can be detected and articulated in the prosaic setting of shopping lists and  recommendations. I will argue that 'conversion' is a preferable term for these changes.  It highlights  re-orientations in subjects, experience, things, numbers and structures that ranges more widely than Hansen's subjective 'understood-to-be-expressed' formulation.

@@

In one of the many industry-meets-academia events occurring in increasingly data science-oriented higher education institutions in the UK, speakers from industry, government and commerce described their work with predictive models.[^11.1] Shreena Patel, a PhD graduate in statistics and operations research,  works as a data scientist for DunnHumby, a well-known [ customer science company ](https://www.dunnhumby.com/). Her work focuses on online grocery shopping at the supermarket chain Tesco.  Speaking to an audience of statisticians and operations researchers, Patel   focused on the models underlying  shopping list recommendations. Her presentation was filled with graphs, numbers, and tables concerning different aspects of the 'Have you forgotten?' recommender system. Patel's presentation was not sociological. In order to explore the operational environment she described, and its predictive operations, I will reconstruct key features of her presentation using sample data drawn from the Tesco API ([Application Programmer Interface](https://devportal.tescolabs.com/docs/services/56c73b1bf205fd0ed81dbe7a/operations/56d5b2bee2813e0a0053c47c), [@Tesco.com_2017])   and a range of data science, statistics and graphics libraries implemented in the widely-used statistical programming language `R` [@RDevelopmentCoreTeam_2010].

@@

@@

Figure \@ref(fig:simulate-inventory) plots something of the density of commodities in the Tesco grocery database. By virtue of their sheer abundance, the plot cannot easily separate the many items that Tesco sells, especially in the price range £1-£10. A sample of the most highly priced items in Tesco stores is shown in Table  \@ref(tab:expensive-groceries)

@@

Patel's presentation presents two opportunities. First, by recounting, contextualising and commenting on some of the main steps in making the 'Have your forgotten?' list  as a example of the predictive sense-making done by data scientists and customer analytics teams working with transactional data in many commercial settings. Patel mentioned many of these steps only fleetingly in the presentation, for they are largely taken for granted as part of predictive analytic practice. What stands out from the presentation is the normality of the recommender system: it is part of a long-standing and ongoing transformation or conversion of grocey shopping. Second, Patel's explanation of the renovation of a recommender system offered its audience a chance to step into the list-making practices situated at the centre of a mundane situation, grocery shopping.  As a sociological response to a data scientist presenting a model to an audience of statisticians and operations researchers, my interest lies less in specific technical innovations and more on how the ordering work done by recommender systems relates to problems of social order more generally.  The method of reconstruction borrows from the philosopher John Dewey a broad interest in how practical and abstract elements of experience are interwoven [@Dewey_1957]. More concretely, it draws heavily on the vision of 'interface methods' proposed by Noortje Marres [@Marres_2015] in order to keep in view both the highly technical orderings of lists in recommender system and the changes in social orderings they might imply or impel.

@@

The main narrative of Patel's presentation concerned a shift from a well-established demographically-targeted data mining model for making recommendations to a predictive, probabilistic, personally relevant  re-writing of the shopping list in almost-realtime. The changes Patel described are increasingly widespread and common. While they are configured in Tesco-specific ways by DunnHumby (and this reflects a longer history), they are also broadly typical. Tesco is the largest supermarket chain in UK, famous for a customer loyalty and targeted marketing programme known as 'Tesco  Clubcard,' which started in 1991. DunnHumby -- founded by operations researchers Edwina Dunn and Clive Humby -- is said to have convinced the CEO of Tesco sometime in 1991 that a loyalty card program could change the supermarket chain's relationship to its customers.  Clive Humby's academic publications  are hard to track down, but an early paper given at the _Conference of Young Operational Researchers_ in Nottingham in 1984 suggests the direction that he, DunnHumby and later Tesco would take in constructing lists [@OKeefe_1984]. The abstract for Humby's presentation pre-figures an ongoing trajectory for data mining techniques aimed at eliciting detailed information on individual customer preferences.

@@

```{r arules-ex, cache=TRUE, message=FALSE, echo=FALSE, fig.margin=TRUE}
library(knitr)
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
options(digits=2)
rules<- apriori(data=Groceries, parameter=list(supp=0.001,conf = 0.15,minlen=2), appearance = list(default="rhs",lhs="whole milk"), control = list(verbose=FALSE))
rules_sorted <- sort(rules, decreasing=TRUE,by="confidence")
kable(inspect(rules_sorted[1:5]))
```

The `apriori` algorithm finds sets of items that commonly occur together in transactions.  These commonly occurring sets are expressed as 'association rules.' For instance, using the `Groceries` dataset in the `R` package `arules,` we can readily see how Walmart *might* have discovered   the fabled association between diapers and beer.[^149.1] Figure \@ref(fig:arules-ex) shows how often common items appear. `Whole milk` appears most frequently.  The `apriori` algorithm attempts to re-structure shopping lists using associations between items frequently bought together.  The association rule-based model for the `Groceries` dataset illustrates some characteristic ordering practices of 1990s-style data-mining approaches to lists (the `Groceries` dataset  was acquired from a 'local German supermarket' [@Hahsler_2006a]). The association between items is expressed in the form of _rules_ not a model. These rules, a term derived from older decision support literature, define sets of associations between items. As the product of a  counting practice, the association rules define sets of items or shopping baskets in terms of decreasing strength. Table \@ref(tab:arules-ex) shows that the first rule has a stronger association than the second. Milk and vegetables is more likely an association than milk and buns.  A rules-based recommender system should, therefore, might put them in that order in the 'Have You Forgotten' list.

@@

```{r tesco-ex, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
    library(ggplot2)
    tesco = read.csv('data/reference_data/tesco_groceries.csv')
    total = sum(tesco$actual, na.rm=TRUE) 
    tesco = na.omit(tesco)
    tesco$actual = as.integer(tesco$actual)
    ggplot(tesco[tesco$actual>50,], aes(x=labels, y=actual, fill=labels)) +geom_bar(stat='identity') + coord_flip() + ggtitle('Tesco grocery items with more than 50 products') + theme(legend.position="none", axis.text=element_text(size=7))
```

Items in the `Groceries` dataset proliferate into a Tesco's list of branded variations. The 169 items of the `Groceries` dataset  expand into roughly `r total` Tesco items (see Figure \@ref(fig:tesco-ex)).  This is, I would suggest, a form of propensity in Hansen's terms, albeit one that occurs in, or very close to, the recommender system rather than  'in the world.' The multiple listed  in the `Grocery` dataset becomes more open in this setting through new impersonal intersections that must be bounded and ordered. The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial. `r total` products (actually Patel mentioned 200,000 products) can be combined ways. If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible lists. Some of this vast number of possible shopping lists are highly unlikely in terms of typical purchasing patterns, but potentially highly interesting and value as recommendation.

@@

```{r sparsity-ex, echo=FALSE, cache=TRUE, fig.margin=FALSE, fig.cap ='Sparsity of the customer-product association matrix' }
library(ggplot2)
library(reshape2)
p = as.logical(rbinom(n=2e5, size=20,0.001 ))
m = (matrix(p, nrow=1000, byrow=TRUE))
melted <- melt(m)
melted$value <- as.logical(melted$value)
ggplot(melted, aes(x = Var2, y = Var1, fill = value)) + geom_tile() +
    scale_fill_manual(values = c("grey", "red")) + xlab('Customer id') + ylab('Product id') + theme(legend.position='none')
```

@@

## Infrastructure: the realtime engine serves

@@

In their intersection with mundane practices of ordering, recommender systems offer some guidance to the ways that predictive practices affect existing forms of social order in the world. Rather than directly positing an ontological transformation, I have explored some of the practices of conversion associated with a shift from a data-mining rule-based system to a predictive recommender system. The DunnHumby model and the Tesco 'Have You Forgotten?' recommendations present many of the traits of the conversion processes associated with big data. Rather than discontinuous changes, the conversion of the recommender system via a relevance model suggests an ongoing problematization of propensities. We have seen several types of conversion events: ongoing transformation of models and infrastructures, the experimentalisation  of recommendations, and the broader diffusion of a calculus of probabilities into a fibrous weave of propensities and conditionalities, warping over time. The predictive assemblages respond to problems of logistical ensembles more generally.

I have been suggesting that predictive listing in the form of open-ended recommendation changes social order. It alters the way in which repetition or reiteration, exemplified in the shopping list, can also participate in change and emergent order.   By following Popper in understanding the world in terms of propensities, we have a chance  to engage with contemporary predictive processes in ways that both acknowledge their relentless and exhausting pressure to consume more and retain some chance of inhabiting the world differently.

I diverge from some understandings of predictive practice. Rather than understanding prediction as the closed form of an ontologically open world ('the propensities of things'), the DunnHumby relevance model, in all its statistical, infrastructural, organisational and banal complexity, suggests that predictive conversion remains constitutively incomplete. Predictive practice abounds in propensities. Ironically, shopping lists  exemplify this incompleteness and ineradicable openness of prediction practice.  The constitutive incompleteness and openness of lists are very closely entangled with problems of social order amidst predictive operations. Lists undergoing a 'conversion' experience as they become more probabilistic, but their heterogeneous character as operations caught up in models, infrastructures, experiments, and the flow of commodities means that they can never form a closed set or dataset. The problem of 'constructing good features,' a problem that any predictive model must address, persists as the site of experiment with structures that lack closure. A constitutive openness to new associations, new combinations, new groupings or actors is integral to the conversion event.

@@

The 'universe of reference' includes objects, ideas and practices mingle. Something similar happens to shopping in the Dunnhumby relevance model. In the customer-product association matrix, in the product-product similarity matrix, in the 14,000 variables included in the logistic regression model, a 'universe of reference' takes shape. The shopping list, the commonplace inscription of normative and actual social order, takes on a complex temporal-material structure that mixes prediction, infrastructural constraints, the individual characteristics of list-makers, statistical validations, database architectures and above all the **surplus sensibility** associated with  circulation of commodities. Shopping lists undergoing a **conversion experience** as they become more probabilistic, but the idea that probabilities are the closed form of propensities does not account for the distribution of probabilities in relations between things, between people, between people and things, and between things and people.
